Disturbance-aware minimum-time planning strategies for motorsport vehicles with probabilistic safety certificates

### 2025-06-11 closed_loop_planning.tex
[MM] typos

\subsection{Step 1 -- Nominal feed-forward time-optimal trajectory planning}
\label{sec:nominalFF}
Step (i) consists of planning a nominal feed-forward trajectory. This assumes deterministic system dynamics, where $\hbmu_k$, $\hbu_k$, and $\hbz_k$ denote differential states, controls, and algebraic states, respectively, for \( k = 0, \ldots, N \). To avoid redundancy, we note that the NLP to be solved is a special case of~\eqref{eq:DOCP}, where~\eqref{eq:DOCPdP},~\eqref{eq:DOCPdPIC}, and the backoff terms are omitted, since the goal is to compute a nominal trajectory.

@@

\subsection{Step 3 -- Closed-loop robustified planning}
\label{sec:CLrobustified}
%\begin{align}
%    \underset{\bmu_k,\bxi_k, \bu_k, \bP_k,\bz_k, \bK_k}{\text{minimize}} \quad & J_k(\bmu_k,\bxi_k, \bu_k) \label{eq:ROCPcost}\\
%       \text{s.t.}\notag\\
%          & f_0(x) - t \le 0 \quad \phantom{i=1, \dots, m}\\
%          & f_i(x) \le 0, \quad i=1, \dots, m \quad i=1, \dots, m\\
%          & h_i(x) = 0,   \quad i=1, \dots, p,
%\end{align}
As the third and last step, a closed-loop robustified planning problem is set up. Its aim is the definition of a mean trajectory of a stochastic system which, working in tandem with a state-feedback controller that aims at stabilizing it, is able to mitigate the perturbations and reduce the propagation of uncertainty. This helps to guarantee that the robustified path constraints can be satisfied over the entire planning horizon. The discretized version of this problem takes the form of the following NLP:
\begin{subequations}\label{eq:ROCP}
\begin{alignat}{3}
\underset{\bmu_k,\bxi_k, \bu_k, \bP_k,\bz_k, \bK_k}{\text{minimize}}
    & \quad J_k(\bmu_k,\bxi_k, \bu_k) & & \label{eq:ROCPcost} \\
\text{s.t.}\notag\\
\bzero      & = \; \bPsimu_k(\bmu_{k-1},\bmu_k,\bxi_k, \bu_k,\bz_k),
\quad k = 1,\ldots, N \label{eq:ROCPdyn} \\
\bmu_0      & = \; \bar{\bmu}_0 \label{eq:ROCPdynIC} \\
\bzero      & = \; \bPsiPCL_k(\bmu_k,\bxi_k, \bu_k, \tbP_{k-1},\tbP_k,\bSi_k,\bz_k, \bK_k),
\quad k = 1,\ldots, N \label{eq:ROCPdP} \\
\tbP_0       & = \; \bar{\bP}_0 \succeq 0 \label{eq:ROCPdPIC} \\
\bzero      & = \; \bOm_k(\bmu_k,\bxi_k, \bu_k,\bz_k), \quad k = 0,\ldots, N \label{eq:ROCPpath} \\
0    & \geq  h_i(\bmu_k, \bu_k, \bz_k) + \be_i(\bmu_k, \bu_k, \tbP_k, \bz_k), \quad k = 1,\ldots, N;\; \quad i \in \calI \label{eq:ROCPconstraints}\\
-\de \bK_k   & \leq \bK_k - \hbK_k \leq  \de \bK_k \label{eq:ROCPdeltaK}, \quad k = 0,\ldots, N
\end{alignat}
\end{subequations}
Here,~\eqref{eq:ROCPdP} denotes the discrete-time covariance dynamics under the stabilizing state-feedback controller \( \bbu_k = -\bK_k \bbmu_k \). To distinguish this case from~\eqref{eq:DOCP}, we use the symbol \( \tbP \) to denote the closed-loop (CL) covariance matrix, whose discrete evolution corresponds to a collocation-based approximation of the continuous-time dynamics $\dot{\tbP}(t) = \tbA(t) \tbP(t) + \tbP(t)\tbA(t)^T + \bQ(t)$, with $\tbA(t)=\bA(t)-\bB(t)\bK(t)$. It is worth noting that, accordingly, in equation~\eqref{eq:ROCPconstraints} the backoff term $\be_i$ is a function of closed-loop uncertainty $\tbP_k$. Another key aspect is that the state-feedback matrix $\bK_k$ is also treated as a decision variable. This reflects the fact that the driver may adjust his action to stabilize the new (to-be-planned) nominal trajectory using a policy that differs from the previous $\hbK_k$, computed in step (ii) via LQR. To promote convergence, however, $\bK_k$ is restrained within prescribed bounds ($\de \bK_k$) relative to $\hbK_k$. In our tests $\de \bK_k = 0.1 \hbK_k$.

>>>>

\subsection{Step 1 -- Nominal feed-forward time-optimal trajectory planning}
\label{sec:nominalFF}
Step (i) consists of planning a nominal feed-forward trajectory. This assumes deterministic system dynamics, where $\hbmu_k$, $\hbu_k$, and $\hbz_k$ denote differential states, controls, and algebraic states, respectively, for \( k = 0, \ldots, N \). To avoid redundancy, we note that the NLP to be solved is a special case of~\eqref{eq:DOCP}, where~\eqref{eq:DOCPdP},~\eqref{eq:DOCPdPIC}, and the back-off terms are omitted, since the goal is to compute a nominal trajectory.

@@

\subsection{Step 3 -- Closed-loop robustified planning}
\label{sec:CLrobustified}
%\begin{align}
%    \underset{\bmu_k,\bxi_k, \bu_k, \bP_k,\bz_k, \bK_k}{\text{minimize}} \quad & J_k(\bmu_k,\bxi_k, \bu_k) \label{eq:ROCPcost}\\
%       \text{s.t.}\notag\\
%          & f_0(x) - t \le 0 \quad \phantom{i=1, \dots, m}\\
%          & f_i(x) \le 0, \quad i=1, \dots, m \quad i=1, \dots, m\\
%          & h_i(x) = 0,   \quad i=1, \dots, p,
%\end{align}
As the third and last step, a closed-loop robustified planning problem is set up. Its aim is the definition of a mean trajectory of a stochastic system which, working in tandem with a state-feedback controller that aims at stabilizing it, is able to mitigate the perturbations and reduce the propagation of uncertainty. This helps to guarantee that the robustified path constraints can be satisfied over the entire planning horizon. The discretized version of this problem takes the form of the following NLP:
\begin{subequations}\label{eq:ROCP}
\begin{alignat}{3}
\underset{\bmu_k,\bxi_k, \bu_k, \bP_k,\bz_k, \bK_k}{\text{minimize}}
    & \quad J_k(\bmu_k,\bxi_k, \bu_k) & & \label{eq:ROCPcost} \\
\text{s.t.}\notag\\
\bzero      & = \; \bPsimu_k(\bmu_{k-1},\bmu_k,\bxi_k, \bu_k,\bz_k),
\quad k = 1,\ldots, N \label{eq:ROCPdyn} \\
\bmu_0      & = \; \bar{\bmu}_0 \label{eq:ROCPdynIC} \\
\bzero      & = \; \bPsiPCL_k(\bmu_k,\bxi_k, \bu_k, \tbP_{k-1},\tbP_k,\bSi_k,\bz_k, \bK_k),
\quad k = 1,\ldots, N \label{eq:ROCPdP} \\
\tbP_0       & = \; \bar{\bP}_0 \succeq 0 \label{eq:ROCPdPIC} \\
\bzero      & = \; \bOm_k(\bmu_k,\bxi_k, \bu_k,\bz_k), \quad k = 0,\ldots, N \label{eq:ROCPpath} \\
0    & \geq  h_i(\bmu_k, \bu_k, \bz_k) + \be_i(\bmu_k, \bu_k, \tbP_k, \bz_k), \quad k = 1,\ldots, N;\; \quad i \in \calI \label{eq:ROCPconstraints}\\
-\de \bK_k   & \leq \bK_k - \hbK_k \leq  \de \bK_k \label{eq:ROCPdeltaK}, \quad k = 0,\ldots, N
\end{alignat}
\end{subequations}
Here,~\eqref{eq:ROCPdP} denotes the discrete-time covariance dynamics under the stabilizing state-feedback controller \( \bbu_k = -\bK_k \bbmu_k \). To distinguish this case from~\eqref{eq:DOCP}, we use the symbol \( \tbP \) to denote the closed-loop (CL) covariance matrix, whose discrete evolution corresponds to a collocation-based approximation of the continuous-time dynamics $\dot{\tbP}(t) = \tbA(t) \tbP(t) + \tbP(t)\tbA(t)^T + \bQ(t)$, with $\tbA(t)=\bA(t)-\bB(t)\bK(t)$. It is worth noting that, accordingly, in equation~\eqref{eq:ROCPconstraints} the back-off term $\be_i$ is a function of closed-loop uncertainty $\tbP_k$. Another key aspect is that the state-feedback matrix $\bK_k$ is also treated as a decision variable. This reflects the fact that the driver may adjust his action to stabilize the new (to-be-planned) nominal trajectory using a policy that differs from the previous $\hbK_k$, computed in step (ii) via LQR. To promote convergence, however, $\bK_k$ is restrained within prescribed bounds ($\de \bK_k$) relative to $\hbK_k$. In our tests $\de \bK_k = 0.1 \hbK_k$.

### 2025-06-11 open_loop_planning.tex
[MGa] curvilinear abscissa -> curvilinear parameter
[MM] typos

The specific formulation is expressed as follows:
\begin{subequations}\label{eq:DOCPopenloop}
\begin{alignat}{3}
	\underset{\bmu_k,\bxi_k, \bu_k, \bP_k,\bz_k}{\text{minimize}} \,
	& & & J_k(\bmu_k,\bxi_k, \bu_k) & & \label{eq:DOCPcostopenloop} \\
	\hspace*{-2.0 cm}\text{s.t.} \quad
	& \bzero      & = & \; \bPsimu_k(\bmu_{k-1},\bmu_k,\bxi_k, \bu_k,\bz_k),
	& \quad & k = 1,\ldots, N \label{eq:DOCPdynopenloop} \\
	& \bmu_0      & = & \; \bar{\bmu}_0
	& & \label{eq:DOCPdynICopenloop} \\
	& \bzero      & = & \; \bPsiP_k(\bmu_k,\bxi_k, \bu_k, \bP_{k-1}^{j-1},\bP_k^j,\bSi_k^j,\bz_k),
	& \quad & \substack{k = 1,\ldots, N; \; \\ j=1,\ldots,H;\;} \label{eq:DOCPdPopenloop} \\
	& \bP_k^0       & = & \; \bar{\bP}_0 \succeq 0
	& \quad & k = 1,\ldots, N; \; \label{eq:DOCPdPinitopenloop} \\
	& \bzero      & = & \; \bOm_k(\bmu_k,\bxi_k, \bu_k,\bz_k),
	& \quad & k = 0,\ldots, N \label{eq:DOCPpathopenloop} \\
	& 0           & \geq & \; h_i(\bmu_k, \bu_k, \bz_k) + \be_i(\bmu_k, \bu_k, \bP_k^H, \bz_k),
	& \quad & k = 1,\ldots, N;\; i \in \calI \label{eq:DOCPconstraintsopenloop}
\end{alignat}
\end{subequations}
Equation~\eqref{eq:DOCPdPinitopenloop} defines the initialization of the appropriate covariance matrix version at step $k$, while Eq.~\eqref{eq:DOCPconstraintsopenloop} specifies that the back-off term in the constraints is computed using the most propagated version, $\bP^H_k$. A schematic illustration of the management of the $H+1$ covariance matrix instances is provided in Figure~\ref{fig:DOCPgrid}. The dashed rectangle highlights the $k$-th discretization step, where the constraints are evaluated. At each grid point, two particular versions of the covariance matrix are emphasized: $\bP^0_k$ (red node), representing the instance to be initialized at that step, and $\bP^H_k$ (green node), corresponding to the version that has been propagated over the full horizon of $H$ steps and is used to robustify the constraints in Eq.~\eqref{eq:DOCPconstraintsopenloop}.

%Eq.~\eqref{eq:DOCPdPinitopenloop} is used to initialize the correct version at the $k$-th step while Eq.~\eqref{eq:DOCPconstraintsopenloop} specify that the back-off term of the constraints is evaluated with the most propagated covariance matrix version, $\bP^H_k$.
%A schematic representation of the approach used to manage the $H+1$ versions of the covariance matrix is shown in Figure \ref{fig:DOCPgrid}, where the dashed rectangle represent the $k$-th step at which the constraints are being formulated. At each grid point, two versions of the covariance matrix are highlighted: $\bP^0_k$ (red node), which is the version to be initialized, and $\bP^H_k$ (green node), which is the most propagated version used to formulate the robust constraints in Eq.~\eqref{eq:DOCPconstraintsopenloop}.
\def\version{marco}
\ifdefstring{\version}{matteo}{
\begin{figure}
	\centering
	\input{Fig/Tikz_Sources/tikz_covariance_evolution_matteo.tex}
	\caption{Schematic representation of the continuity and initialization equations associated with the $H+1$ propagated instances of the covariance matrix introduced in problem~\eqref{eq:DOCPopenloop}. The dashed rectangle highlights the $k$-th discretization step, at which the robust constraints are enforced. Each node corresponds to a specific covariance matrix, where the subscript denotes the current grid index and the superscript indicates the number of propagation steps since initialization. At each step, two key instances are emphasized: the matrix to be initialized at that step (red node), and the matrix that has been propagated over $H$ steps (green node), which is used in the evaluation of the constraint backoff.}
	%\caption{Schematic representation of the continuity and initialization Equations for the $H+1$ versions of the covariance matrix introduced in the discretized OCP. The dashed rectangle indicate the $k$-th step, in which the constraints are formulated. Each node represents a covariance matrix, labeled such that the subscript denotes the current grid step, while the superscript indicates the number of propagation steps it has undergone. At each step, two nodes are highlighted: the covariance matrix to be initialized (red node) and the matrix that has been propagated for $H$ steps (green node), which is used to formulate the robust constraints.}
	\label{fig:DOCPgrid}
\end{figure}
}{}

\ifdefstring{\version}{marco}{
\begin{figure}
	\centering
	\definecolor{mygreen}{RGB}{0,150,0}
	\input{Fig/Tikz_Sources/tikz_covariance_evolution.tex}
	\caption{Schematic representation (depicted for $H=3$) of the continuity and initialization equations associated with the $H+1$ propagated instances of the covariance matrix introduced in problem~\eqref{eq:DOCPopenloop}. The dashed rectangle highlights the $k$-th discretization step, at which the robust constraints are enforced. Each node corresponds to a specific covariance matrix, where the subscript denotes the current grid index and the superscript indicates the number of propagation steps since initialization. At each step, two key instances are emphasized: the matrix to be initialized at that step (red node), and the matrix that has been propagated over $H$ steps (green node), which is used in the evaluation of the constraint backoff.}
	%\caption{Schematic representation of the continuity and initialization Equations for the $H+1$ versions of the covariance matrix introduced in the discretized OCP. The dashed rectangle indicate the $k$-th step, in which the constraints are formulated. Each node represents a covariance matrix, labeled such that the subscript denotes the current grid step, while the superscript indicates the number of propagation steps it has undergone. At each step, two nodes are highlighted: the covariance matrix to be initialized (red node) and the matrix that has been propagated for $H$ steps (green node), which is used to formulate the robust constraints.}
	\label{fig:DOCPgrid}
\end{figure}
}{}

>>>>

The specific formulation is expressed as follows:
\begin{subequations}\label{eq:DOCPopenloop}
\begin{alignat}{3}
	\underset{\bmu_k,\bxi_k, \bu_k, \bP_k,\bz_k}{\text{minimize}} \,
	& & & J_k(\bmu_k,\bxi_k, \bu_k) & & \label{eq:DOCPcostopenloop} \\
	\hspace*{-2.0 cm}\text{s.t.} \quad
	& \bzero      & = & \; \bPsimu_k(\bmu_{k-1},\bmu_k,\bxi_k, \bu_k,\bz_k),
	& \quad & k = 1,\ldots, N \label{eq:DOCPdynopenloop} \\
	& \bmu_0      & = & \; \bar{\bmu}_0
	& & \label{eq:DOCPdynICopenloop} \\
	& \bzero      & = & \; \bPsiP_k(\bmu_k,\bxi_k, \bu_k, \bP_{k-1}^{j-1},\bP_k^j,\bSi_k^j,\bz_k),
	& \quad & \substack{k = 1,\ldots, N; \; \\ j=1,\ldots,H;\;} \label{eq:DOCPdPopenloop} \\
	& \bP_k^0       & = & \; \bar{\bP}_0 \succeq 0
	& \quad & k = 1,\ldots, N; \; \label{eq:DOCPdPinitopenloop} \\
	& \bzero      & = & \; \bOm_k(\bmu_k,\bxi_k, \bu_k,\bz_k),
	& \quad & k = 0,\ldots, N \label{eq:DOCPpathopenloop} \\
	& 0           & \geq & \; h_i(\bmu_k, \bu_k, \bz_k) + \be_i(\bmu_k, \bu_k, \bP_k^H, \bz_k),
	& \quad & k = 1,\ldots, N;\; i \in \calI \label{eq:DOCPconstraintsopenloop}
\end{alignat}
\end{subequations}
Equation~\eqref{eq:DOCPdPinitopenloop} defines the initialization of the appropriate covariance matrix version at step $k$, while Eq.~\eqref{eq:DOCPconstraintsopenloop} specifies that the back-off term in the constraints is computed using the most propagated version, $\bP^H_k$. A schematic illustration of the management of the $H+1$ covariance matrix instances is provided in Fig.~\ref{fig:DOCPgrid}. The dashed rectangle highlights the $k$-th discretization step, where the constraints are evaluated. At each grid point, two particular versions of the covariance matrix are emphasized: $\bP^0_k$ (red node), representing the instance to be initialized at that step, and $\bP^H_k$ (green node), corresponding to the version that has been propagated over the full horizon of $H$ steps and is used to robustify the constraints in Eq.~\eqref{eq:DOCPconstraintsopenloop}.

%Eq.~\eqref{eq:DOCPdPinitopenloop} is used to initialize the correct version at the $k$-th step while Eq.~\eqref{eq:DOCPconstraintsopenloop} specify that the back-off term of the constraints is evaluated with the most propagated covariance matrix version, $\bP^H_k$.
%A schematic representation of the approach used to manage the $H+1$ versions of the covariance matrix is shown in Figure \ref{fig:DOCPgrid}, where the dashed rectangle represent the $k$-th step at which the constraints are being formulated. At each grid point, two versions of the covariance matrix are highlighted: $\bP^0_k$ (red node), which is the version to be initialized, and $\bP^H_k$ (green node), which is the most propagated version used to formulate the robust constraints in Eq.~\eqref{eq:DOCPconstraintsopenloop}.
\def\version{marco}
\ifdefstring{\version}{matteo}{
\begin{figure}
	\centering
	\input{Fig/Tikz_Sources/tikz_covariance_evolution_matteo.tex}
	\caption{Schematic representation of the continuity and initialization equations associated with the $H+1$ propagated instances of the covariance matrix introduced in problem~\eqref{eq:DOCPopenloop}. The dashed rectangle highlights the $k$-th discretization step, at which the robust constraints are enforced. Each node corresponds to a specific covariance matrix, where the subscript denotes the current grid index and the superscript indicates the number of propagation steps since initialization. At each step, two key instances are emphasized: the matrix to be initialized at that step (red node), and the matrix that has been propagated over $H$ steps (green node), which is used in the evaluation of the constraint back-off.}
	%\caption{Schematic representation of the continuity and initialization Equations for the $H+1$ versions of the covariance matrix introduced in the discretized OCP. The dashed rectangle indicate the $k$-th step, in which the constraints are formulated. Each node represents a covariance matrix, labeled such that the subscript denotes the current grid step, while the superscript indicates the number of propagation steps it has undergone. At each step, two nodes are highlighted: the covariance matrix to be initialized (red node) and the matrix that has been propagated for $H$ steps (green node), which is used to formulate the robust constraints.}
	\label{fig:DOCPgrid}
\end{figure}
}{}

\ifdefstring{\version}{marco}{
\begin{figure}
	\centering
	\definecolor{mygreen}{RGB}{0,150,0}
	\input{Fig/Tikz_Sources/tikz_covariance_evolution.tex}
	\caption{Schematic representation (depicted for $H=3$) of the continuity and initialization equations associated with the $H+1$ propagated instances of the covariance matrix introduced in problem~\eqref{eq:DOCPopenloop}. The dashed rectangle highlights the $k$-th discretization step, at which the robust constraints are enforced. Each node corresponds to a specific covariance matrix, where the subscript denotes the current grid index and the superscript indicates the number of propagation steps since initialization. At each step, two key instances are emphasized: the matrix to be initialized at that step (red node), and the matrix that has been propagated over $H$ steps (green node), which is used in the evaluation of the constraint back-off.}
	%\caption{Schematic representation of the continuity and initialization Equations for the $H+1$ versions of the covariance matrix introduced in the discretized OCP. The dashed rectangle indicate the $k$-th step, in which the constraints are formulated. Each node represents a covariance matrix, labeled such that the subscript denotes the current grid step, while the superscript indicates the number of propagation steps it has undergone. At each step, two nodes are highlighted: the covariance matrix to be initialized (red node) and the matrix that has been propagated for $H$ steps (green node), which is used to formulate the robust constraints.}
	\label{fig:DOCPgrid}
\end{figure}
}{}

### 2025-06-11 stochastic_vehicle_dynamics.tex
[MM] typos

As can be readily verified by differentiation~\cite{Gajic:LyapunovMatrixEquation:2010}, the analytical solution of~\eqref{eq:dP} has the form
\begin{align}\label{eq:P_STM}
\bP(t) = \bPhi(t,t_0) \bP_0 \bPhi(t,t_0)+\int_{t_0}^{t} \bPhi(t,\tau) \bQ(\tau) \bPhi^T(t,\tau) \dd \tau \quad \bP(0)=\bP_0,
\end{align}
where $\bPhi(t,t_0)$ is the \emph{state transition matrix}. This matrix encodes the evolution from $t_0$ to $t$ of a perturbation w.r.t. to the mean trajectory $\bmu(t)$. In symbols, $\barbx(t) = \bPhi(t,t_0) \barbx(t_0)$, with $\barbx(\cdot) =\bx(\cdot) - \bmu(\cdot) $. In turn, the evolution of $\bPhi(t,t_k)$ from a generic $t_k$ to $t$ is driven by the following differential equation
\begin{align}\label{eq:STM}
\dbPhi(t,t_k) = \bA(t)\bPhi(t,t_k), \quad \bPhi(t_k, t_k) = \bI.
\end{align}
In geometric and orbital mechanics, see e.g.~\cite{Maruskin:DynamicalSystemsGeometric:2018} or~\cite{Tapley:StatisticalOrbitDetermination:2004}, the usual choice for statistical trajectory determination is to employ~\eqref{eq:meandynamics} along with~\eqref{eq:P_STM} and~\eqref{eq:STM}. In our case, since we use collocation integrators for stochastic trajectory planning, we follow a more direct approach -- motivated by~\cite{Gillis:PracticalMethodsApproximate:2015} -- by directly employing~\eqref{eq:meandynamics} and~\eqref{eq:dP}.

Accordingly, the continuous-time stochastic trajectory planning can be framed as the following nonlinear optimal control problem
\begin{subequations}\label{eq:OCP}
\begin{align}
	\underset{\bmu(t), \bu(t), \bP(t)}{\text{minimize}} \quad & J(\bmu(t), \bu(t), \bP(t)) \label{eq:OCPcost} \\
	\text{s.t.} \quad \dbmu(t)           &= f(\bmu(t), \bu(t)) \label{eq:OCPdyn} \\
	\phantom{\text{s.t.} \quad} \bmu(0)  &= \bmu_0 \label{eq:OCPdynIC} \\
	\phantom{\text{s.t.} \quad} \dbP(t) &= \bA(t) \bP(t) + \bP(t) \bA^T(t) + \bQ(t) \label{eq:OCPdP} \\
	\phantom{\text{s.t.} \quad} \bP(0)  &= \bP_0 \succeq 0 \label{eq:OCPdPIC} \\ %& \phantom{\text{s.t.} \qquad} 0       \geq h_i(\bmu(t), \bu(t))
	%+ \overbrace{\gamma \bigg[\underbrace{\na^T_{\bx} h_i \bP(t) \na_{\bx} h_i}_{\text{variance of $h_i(\bx)$}}\bigg]^{\frac{1}{2}}}^{\text{safety margin}},
	\phantom{\text{s.t.} \qquad} 0&       \geq h_i(\bmu(t), \bu(t))
	+ \be_i(\bmu(t), \bu(t), \bP(t)),
	\quad i \in \calI \label{eq:OCPconstraints}
\end{align}
\end{subequations}
The cost function $J$ in~\eqref{eq:OCPcost} depends on the mean $\bmu(t)$, the controls $\bu(t)$ and the state covariance $\bP(t)$. The mean dynamics is expressed by~\eqref{eq:OCPdyn} with~\eqref{eq:OCPdynIC}, and the covariance dynamics is expressed by~\eqref{eq:OCPdP} with~\eqref{eq:OCPdPIC}. In eq.~\eqref{eq:OCPconstraints} the \emph{backoff terms} $\be_i$ account for the disturbances and serve the purpose of obtaining deterministic safety margins directly on $\bmu(t)$. $\calI$ is the set of indices defining the inequality constraints. The backoff terms stem for a linearization around the mean of the original \emph{chance constraint} on $\bx(t)$ expressed by $\Prob \{h_i(\bx) \leq 0\}\geq p$, where $p$ is the confidence level of constraint satisfaction. Explicitly, the backoff terms are given by
$\be_i = \gamma \sig_i$. The coefficient $\gamma = \Phi^{-1}(p)$ is the quantile function, where $\Phi(z)=\Prob \{Z\leq z\}$ is the cumulative distribution function (cdf) of a standard normal distribution $Z \sim \calN(0,1)$, and acts as a \emph{tuning knob}: the greater the confidence level $p$ required, the higher the gain $\ga$\footnote{For example, with $p=0.84$ $\ga = 1.0$, with $p=0.97$ $\ga = 2.0$, with $p=0.99$ $\ga = 3.0$}. The term $\sig_i = \big[\na^T_{\bx} h_i(\bmu) \bP(t) \na_{\bx} h_i(\bmu)\big]^{\frac{1}{2}}$ represents the standard deviation of the constraint $h_i(\bx)$ linearized around the mean, i.e. of random variable $h_i(\bmu)+\na^T_{\bx} h_i(\bmu) (\bx - \bmu)$, and follows immediately from the propagation rule of covariance.

@@

\subsection{Robust track limit constraint formulation}
\label{sec:TLC}
An track limit constraint is required to ensure that the vehicle remains within the circuit boundaries. The formulation adopted introduces an algebraic variable $e$, which satisfies the following constraint:
\begin{equation}
	\begin{Bmatrix}
		x\\y
	\end{Bmatrix}
	-\boldsymbol{c} - e\bn=\boldsymbol{0} \label{eq:onplane_constraint}
\end{equation}
%\begin{equation}
%	\bp-\boldsymbol{c}_k - e_k\bn_k=\boldsymbol{0} \label{eq:onplane_constraint}
%\end{equation}
where $\boldsymbol{c}$ indicates the position of the centerline, and $\bn$ is the unit normal vector to the centerline. Equation~\eqref{eq:onplane_constraint} constraints the CoM of the vehicle to lie on the vertical plane defined by the unit normal vector $\bn$. The previously introduced algebraic variable $e$ is further constrained to take values between $e_\textrm{min}$ and $e_\textrm{max}$, which are determined based on the width of the circuit and the vehicle's front and rear tracks.

In order to take into account for the back-off terms, the maximum and minimum values are modified coherently, so that:
\begin{equation}
	e_\textrm{min} + \be^\textrm{TLC} \leq e \leq e_\textrm{max} - \be^\textrm{TLC}\label{eq:TLC}
\end{equation}
where $\be^\textrm{TLC}$ is computed by considering that the gradient of the constraint w.r.t. the state vector $\na_{\bx}h^\textrm{TLC}$ depends solely on $\bn$.
%Adding a back-off to this constraint produces the same effect as an increase of the front and rear tracks of the vehicle does.

>>>>

As can be readily verified by differentiation~\cite{Gajic:LyapunovMatrixEquation:2010}, the analytical solution of~\eqref{eq:dP} has the form
\begin{align}\label{eq:P_STM}
\bP(t) = \bPhi(t,t_0) \bP_0 \bPhi^T(t,t_0)+\int_{t_0}^{t} \bPhi(t,\tau) \bQ(\tau) \bPhi^T(t,\tau) \dd \tau \quad \bP(0)=\bP_0,
\end{align}
where $\bPhi(t,t_0)$ is the \emph{state transition matrix}. This matrix encodes the evolution from $t_0$ to $t$ of a perturbation w.r.t. to the mean trajectory $\bmu(t)$. In symbols, $\barbx(t) = \bPhi(t,t_0) \barbx(t_0)$, with $\barbx(\cdot) =\bx(\cdot) - \bmu(\cdot) $. In turn, the evolution of $\bPhi(t,t_k)$ from a generic $t_k$ to $t$ is driven by the following differential equation
\begin{align}\label{eq:STM}
\dbPhi(t,t_k) = \bA(t)\bPhi(t,t_k), \quad \bPhi(t_k, t_k) = \bI.
\end{align}
In geometric and orbital mechanics, see e.g.~\cite{Maruskin:DynamicalSystemsGeometric:2018} or~\cite{Tapley:StatisticalOrbitDetermination:2004}, the usual choice for statistical trajectory determination is to employ~\eqref{eq:meandynamics} along with~\eqref{eq:P_STM} and~\eqref{eq:STM}. In our case, since we use collocation integrators for stochastic trajectory planning, we follow a more direct approach -- motivated by~\cite{Gillis:PracticalMethodsApproximate:2015} -- by directly employing~\eqref{eq:meandynamics} and~\eqref{eq:dP}.

Accordingly, the continuous-time stochastic trajectory planning can be framed as the following nonlinear optimal control problem
\begin{subequations}\label{eq:OCP}
\begin{align}
	\underset{\bmu(t), \bu(t), \bP(t)}{\text{minimize}} \quad & J(\bmu(t), \bu(t), \bP(t)) \label{eq:OCPcost} \\
	\text{s.t.} \quad \dbmu(t)           &= f(\bmu(t), \bu(t)) \label{eq:OCPdyn} \\
	\phantom{\text{s.t.} \quad} \bmu(0)  &= \bmu_0 \label{eq:OCPdynIC} \\
	\phantom{\text{s.t.} \quad} \dbP(t) &= \bA(t) \bP(t) + \bP(t) \bA^T(t) + \bQ(t) \label{eq:OCPdP} \\
	\phantom{\text{s.t.} \quad} \bP(0)  &= \bP_0 \succeq 0 \label{eq:OCPdPIC} \\ %& \phantom{\text{s.t.} \qquad} 0       \geq h_i(\bmu(t), \bu(t))
	%+ \overbrace{\gamma \bigg[\underbrace{\na^T_{\bx} h_i \bP(t) \na_{\bx} h_i}_{\text{variance of $h_i(\bx)$}}\bigg]^{\frac{1}{2}}}^{\text{safety margin}},
	\phantom{\text{s.t.} \qquad} 0&       \geq h_i(\bmu(t), \bu(t))
	+ \be_i(\bmu(t), \bu(t), \bP(t)),
	\quad i \in \calI \label{eq:OCPconstraints}
\end{align}
\end{subequations}
The cost function $J$ in~\eqref{eq:OCPcost} depends on the mean $\bmu(t)$, the controls $\bu(t)$ and the state covariance $\bP(t)$. The mean dynamics is expressed by~\eqref{eq:OCPdyn} with~\eqref{eq:OCPdynIC}, and the covariance dynamics is expressed by~\eqref{eq:OCPdP} with~\eqref{eq:OCPdPIC}. In eq.~\eqref{eq:OCPconstraints} the \emph{back-off terms} $\be_i$ account for the disturbances and serve the purpose of obtaining deterministic safety margins directly on $\bmu(t)$. $\calI$ is the set of indices defining the inequality constraints. The back-off terms stem from a linearization around the mean of the original \emph{chance constraint} on $\bx(t)$ expressed by $\Prob \{h_i(\bx) \leq 0\}\geq p$, where $p$ is the confidence level of constraint satisfaction. Explicitly, the back-off terms are given by
$\be_i = \gamma \sig_i$. The coefficient $\gamma = \Phi^{-1}(p)$ is the quantile function, where $\Phi(z)=\Prob \{Z\leq z\}$ is the cumulative distribution function (cdf) of a standard normal distribution $Z \sim \calN(0,1)$, and acts as a \emph{tuning knob}: the greater the confidence level $p$ required, the higher the gain $\ga$\footnote{For example, with $p=0.84$ $\ga = 1.0$, with $p=0.97$ $\ga = 2.0$, with $p=0.99$ $\ga = 3.0$}. The term $\sig_i = \big[\na^T_{\bx} h_i(\bmu) \bP(t) \na_{\bx} h_i(\bmu)\big]^{\frac{1}{2}}$ represents the standard deviation of the constraint $h_i(\bx)$ linearized around the mean, i.e. of random variable $h_i(\bmu)+\na^T_{\bx} h_i(\bmu) (\bx - \bmu)$, and follows immediately from the propagation rule of covariance.

@@

\subsection{Robust track limit constraint formulation}
\label{sec:TLC}
A track limit constraint is required to ensure that the vehicle remains within the circuit boundaries. The formulation adopted introduces an algebraic variable $e$, which satisfies the following constraint:
\begin{equation}
	\begin{Bmatrix}
		x\\y
	\end{Bmatrix}
	-\boldsymbol{c} - e\bn=\boldsymbol{0} \label{eq:onplane_constraint}
\end{equation}
%\begin{equation}
%	\bp-\boldsymbol{c}_k - e_k\bn_k=\boldsymbol{0} \label{eq:onplane_constraint}
%\end{equation}
where $\boldsymbol{c}$ indicates the position of the centerline, and $\bn$ is the unit normal vector to the centerline. Equation~\eqref{eq:onplane_constraint} constraints the CoM of the vehicle to lie on the vertical plane defined by the unit normal vector $\bn$. The previously introduced algebraic variable $e$ is further constrained to take values between $e_\textrm{min}$ and $e_\textrm{max}$, which are determined based on the width of the circuit and the vehicle's front and rear tracks.

In order to take into account the back-off terms, the maximum and minimum values are modified coherently, so that:
\begin{equation}
	e_\textrm{min} + \be^\textrm{TLC} \leq e \leq e_\textrm{max} - \be^\textrm{TLC}\label{eq:TLC}
\end{equation}
where $\be^\textrm{TLC}$ is computed by considering that the gradient of the constraint w.r.t. the state vector $\na_{\bx}h^\textrm{TLC}$ depends solely on $\bn$.
%Adding a back-off to this constraint produces the same effect as an increase of the front and rear tracks of the vehicle does.
